---
title: "Order Flow Analysis"
author: "N.Naidoo, D. Chetty"
format: pdf
---
# Setup for analysis

## Required libraries
```{r}
library(dplyr) # pipe operator
library(stringr) # string replace
library(lubridate)
library(ggplot2) #plots
library(zoo) # resampling if needed
library(tseries) # ADF + KPSS tests
library(tidyr)
library(knitr) # display tables
library(data.table)
library(lubridate)
library(urca)
library(patchwork)
```


## Loading in data
```{r}
quote_data_all<-read.csv("XBTUSD_quotes_20171001_20171023.csv",header = TRUE)
trade_data_all<-read.csv("XBTUSD_trades_20171001_20171023.csv",header = TRUE)
# save(quote_data_all,trade_data_all, file = "trade_and_quote_data.RData")
# load("trade_and_quote_data.RData") For REMOTE Use only as file is too big to put on github so do not commit.
```

## Cleaning and formatting data 
```{r}
## Cleaning data so it matches data description of Silyantyev(2019)
quote_data_all<-quote_data_all |> select(timestamp,bidPrice,bidSize,askPrice,askSize) 
quote_data_all<-quote_data_all |> rename(bidVolume=bidSize,askVolume=askSize)

## Removing D character in csv files for timestamp column so it can be parsed correctly
quote_data_all <- quote_data_all |>
  mutate(
    timestamp = str_replace(timestamp, "D", " "),
  )

trade_data_all<-trade_data_all |> select(timestamp,price,size,side)
trade_data_all<-trade_data_all |> rename(Volume=size)
trade_data_all <- trade_data_all |>
  mutate(
    timestamp = str_replace(timestamp, "D", " "),
  )

```

## Quote data: Additional columns - timestamp,midprice and OFI column 
```{r}

## Calculating mid price
## Format timestamp so it can be parsed correctly,ensuring no loss of information
quote_data_all <- quote_data_all |>
  mutate(
    timestamp = as.POSIXct(timestamp, format = "%Y-%m-%d %H:%M:%OS", tz = "UTC"),
    mid_price = (bidPrice + askPrice) / 2
  )

# Calculating OFI 
quote_data_all <- quote_data_all |>
  arrange(timestamp) |>
  mutate(
    prev_bidPrice = lag(bidPrice),
    prev_bidVolume = lag(bidVolume),
    prev_askPrice = lag(askPrice),
    prev_askVolume = lag(askVolume),

    bid_geq = bidPrice >= prev_bidPrice,
    bid_leq = bidPrice <= prev_bidPrice,
    ask_geq = askPrice >= prev_askPrice,
    ask_leq = askPrice <= prev_askPrice,

    OFI = 0
  ) |>
  mutate(
    OFI = OFI +
      if_else(bid_geq, bidVolume, 0) -
      if_else(bid_leq, prev_bidVolume, 0) +
      if_else(ask_geq, prev_askVolume, 0) -
      if_else(ask_leq, askVolume, 0)
  )

```


## Trade data: Additional columns- timestamp,signed_volume 
```{r}
#Format trade data for tfi
trade_data_all <- trade_data_all  |> 
  mutate(
    timestamp = as.POSIXct(timestamp, format = "%Y-%m-%d %H:%M:%OS", tz = "UTC"),
    signed_volume = if_else(side == "Buy", Volume, -Volume) 
  )


```

## Functions for OFI vs Mid-price and TFI vs Mid-price scatter plots
```{r}
# Function to Aggregate to Intervals
# Resample quotes to get mid-price change and OFI per interval

aggregate_ofi <- function(df, interval = "10 secs", tick_size = 0.1) {
  df |>
    mutate(interval_time = floor_date(timestamp, unit = interval)) |>
    group_by(interval_time) |>
    summarise(
      mid_price_start = first(mid_price),
      mid_price_end = last(mid_price),
      OFI = sum(OFI, na.rm = TRUE)
    ) |>
    mutate(mid_price_change = (mid_price_end - lag(mid_price_end)) / tick_size) |>
    drop_na()
}

# Resample trades to get TFI per interval
aggregate_tfi <- function(trades, quotes, interval = "10 secs", tick_size = 0.1) {
  tfi <- trades |>
    mutate(interval_time = floor_date(timestamp, unit = interval)) |>
    group_by(interval_time) |>
    summarise(TFI = sum(signed_volume, na.rm = TRUE))

  mid <- quotes |>
    mutate(interval_time = floor_date(timestamp, unit = interval)) |>
    group_by(interval_time) |>
    summarise(mid_price = last(mid_price)) |>
    mutate(mid_price_change = (mid_price - lag(mid_price)) / tick_size)

  inner_join(mid, tfi, by = "interval_time") |> drop_na()
}

```


## Function for OFI and TFI Regression Analysis
```{r}
# function to run lm an extraxt relevant stats
run_ols <- function(df, xvar, yvar) {
  formula <- as.formula(paste(yvar, "~", xvar))
  model <- lm(formula, data = df)
  s <- summary(model)
  c(
    alpha = s$coefficients[1, 1],
    beta = s$coefficients[2, 1],
    t_beta = s$coefficients[2, 3],
    r2 = s$r.squared,
    sig = (1 - s$coefficients[2, 4]) * 100
  )
}
```



# 5.1  Statistical properties
## 5.1.1 Prices

### Figure 1: Mid-price over time (Price chart)
```{r}
#| fig-width: 9
#| fig-height: 3
# Mid-price = (Bid price + Ask price) / 2
# Plot mid-price over time for the selected day.
# 15th October 2017
# Prepare mid-price data for 15 Oct 2017
quote_data_day <- quote_data_all |>
  mutate(timestamp = ymd_hms(timestamp, tz = "UTC")) |>
  filter(as_date(timestamp) == as_date("2017-10-15")) |>
  mutate(mid_price = (bidPrice + askPrice) / 2)

start_time <- as.POSIXct("2017-10-15 03:00:00", tz = "UTC")
end_time   <- as.POSIXct("2017-10-15 21:00:00", tz = "UTC")
breaks_seq <- seq(from = start_time, to = end_time, by = "3 hours")

# XBTUSD Mid-Price Series (15 Oct 2017)
ggplot(quote_data_day, aes(x = timestamp, y = mid_price, color = "midprice")) +
  geom_line() +
  scale_color_manual(values = c("midprice" = "blue")) +
  scale_x_datetime(
    breaks = breaks_seq,
    labels = format(breaks_seq, "%m.%d.%H", tz = "UTC"),
    timezone = "UTC"
  ) +
  labs(
    title = "",
    x = "", y = "Mid-price ($)",
    color = NULL
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
     axis.title.x = element_text(margin = margin(t = 10)),
  axis.title.y = element_text(margin = margin(r = 10)),  
    panel.grid.major = element_line(color = "gray80"),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    legend.position = c(1, 1),
    legend.justification = c(1, 1),
    legend.background = element_rect(fill = "white", color = "black"),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.3)
  )


# Mid-price plot: shows the price movements over the trading day

```

### Figure 1: Change in Mid-price over time 
```{r}
#| fig-width: 9
#| fig-height: 3
# Mid-price change = current mid-price − previous mid-price.
# This gives tick-to-tick price movements.
# Plot the differenced series over time
# Ensure timestamps are in UTC
quote_data_day$timestamp <- as.POSIXct(quote_data_day$timestamp, tz = "UTC")

# Calculate tick-to-tick mid-price change
quote_data_day <- quote_data_day |>
  arrange(timestamp) |>
  mutate(mid_price_change = c(NA, diff(mid_price)))

# Define 3-hour breaks in UTC
start_time <- as.POSIXct("2017-10-15 03:00:00", tz = "UTC")
end_time   <- as.POSIXct("2017-10-15 21:00:00", tz = "UTC")
breaks_seq <- seq(from = start_time, to = end_time, by = "3 hours")

# Plot # Tick-to-Tick Mid-Price Change (15 Oct 2017)
ggplot(quote_data_day, aes(x = timestamp, y = mid_price_change, color = "dmidprice")) +
  geom_line() +
  scale_color_manual(values = c("dmidprice" = "orange")) +
  scale_x_datetime(
    breaks = breaks_seq,
    labels = format(breaks_seq, "%m.%d.%H", tz = "UTC"),
    timezone = "UTC"
  ) +
  labs(
    title = "",
    x = "", y = bquote(Delta ~ "Mid-price ($)"),
    color = NULL
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.x = element_text(margin = margin(t = 10)),
  axis.title.y = element_text(margin = margin(r = 10)),  
    panel.grid.major = element_line(color = "gray80"),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    legend.position = c(1, 1),
    legend.justification = c(1, 1),
    legend.background = element_rect(fill = "white", color = "black"),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.3)
  )

```

### Figure 1: Rolling volatility of mid-price changes 
```{r}
#| fig-width: 9
#| fig-height: 3
# Proxy for volatility: moving standard deviation of mid-price changes.
# Rolling window: 1000 ticks, as per the paper.

# Ensure timestamps are in UTC
quote_data_day$timestamp <- as.POSIXct(quote_data_day$timestamp, tz = "UTC")

# Rolling volatility
quote_data_day$rolling_volatility <- rollapply(
  quote_data_day$mid_price_change,
  width = 1000,
  FUN = sd,
  fill = NA,
  align = "right"
)

# Breaks with UTC
start_time <- as.POSIXct("2017-10-15 03:00:00", tz = "UTC")
end_time   <- as.POSIXct("2017-10-15 21:00:00", tz = "UTC")
breaks_seq <- seq(from = start_time, to = end_time, by = "3 hours")

# Plot # Rolling Volatility (1000-tick window)
ggplot(quote_data_day, aes(x = timestamp, y = rolling_volatility, color = "vol")) +
  geom_line() +
  scale_color_manual(values = c("vol" = "darkgreen")) +
  scale_x_datetime(
    breaks = breaks_seq,
    labels = format(breaks_seq, "%m.%d.%H", tz = "UTC"),
    timezone = "UTC"
  ) +
  labs(
    title = "",
    x = "", y = "Rolling volatility",
    color = NULL
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.x = element_text(margin = margin(t = 10)),
  axis.title.y = element_text(margin = margin(r = 10)),  
    panel.grid.major = element_line(color = "gray80"),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    legend.position = c(1, 1),
    legend.justification = c(1, 1),
    legend.background = element_rect(fill = "white", color = "black"),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.3)
  )

# Rolling volatility plot: shows estimated volatility over the day

```



### Figure 1: Plots vertically stacked
```{r}
# Define common breaks and labels
start_time <- as.POSIXct("2017-10-15 03:00:00", tz = "UTC")
end_time   <- as.POSIXct("2017-10-15 21:00:00", tz = "UTC")
breaks_seq <- seq(from = start_time, to = end_time, by = "3 hours")
labels_seq <- format(breaks_seq, "%m.%d.%H", tz = "UTC")

# --- Plot 1: Mid-Price ---
p1 <- ggplot(quote_data_day, aes(x = timestamp, y = mid_price, color = "midprice")) +
  geom_line() +
  scale_color_manual(values = c("midprice" = "blue")) +
  scale_x_datetime(breaks = breaks_seq, labels = NULL) +
  labs(x = NULL, y = "Mid-price ($)", color = NULL) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_text(size = 10),   # y-axis title
    panel.grid.major = element_line(color = "gray80"),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.3),
    legend.position = c(1, 1),
    legend.justification = c(1, 1),
    legend.background = element_rect(fill = "white", color="black"),
    legend.text  = element_text(size = 8),    # legend labels
  legend.title = element_text(size = 9)     # legend title (if present)
  )

# --- Plot 2: ΔMid-Price ---
p2 <- ggplot(quote_data_day, aes(x = timestamp, y = mid_price_change, color = "dmidprice")) +
  geom_line() +
  scale_color_manual(values = c("dmidprice" = "orange")) +
  scale_x_datetime(breaks = breaks_seq, labels = NULL) +
  labs(
    x = NULL,
    y = bquote(Delta ~ "Mid-price ($)"),  # Greek small delta
    color = NULL
  ) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_text(size = 10),   # y-axis title
    panel.grid.major = element_line(color = "gray80"),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.3),
    legend.position = c(1, 1),
    legend.justification = c(1, 1),
    legend.background = element_rect(fill = "white", color = "black"),
    legend.text  = element_text(size = 8),    # legend labels
  legend.title = element_text(size = 9)     # legend title (if present)
  )

# --- Plot 3: Rolling Volatility ---
p3 <- ggplot(quote_data_day, aes(x = timestamp, y = rolling_volatility, color = "vol")) +
  geom_line() +
  scale_color_manual(values = c("vol" = "darkgreen")) +
  scale_x_datetime(breaks = breaks_seq, labels = labels_seq) +
  labs(x = "", y = "Rolling volatility", color = NULL) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.y = element_text(size = 10),   # y-axis title
    panel.grid.major = element_line(color = "gray80"),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.3),
    legend.position = c(1, 1),
    legend.justification = c(1, 1),
    legend.background = element_rect(fill = "white", color = "black"),
    legend.text  = element_text(size = 8),    # legend labels
  legend.title = element_text(size = 9)     # legend title (if present)
  )

# --- Combine vertically ---
p1 / p2 / p3


```


### Dickey-Fuller and KPSS test of stationarity for Price series 
```{r}
#To examine stationarity of the differenced price series, augmented Dickey–Fuller (ADF) and KPSS tests are conducted or all variants of k in ΔPk variable. The tests confirm that price series are stationary  for every sampling period k at 1% significance level.
library(dplyr)
library(lubridate)
library(urca)
library(knitr)

intervals <- c("1 sec", "10 sec","1 min", "5 min", "10 min", "1 hour")

perform_stationarity_urca <- function(series) {
  series <- as.numeric(series)
  series <- series[is.finite(series)]
  
  if (length(series) < 10) {
    return(data.frame(
      ADF_Statistic = NA_real_, ADF_CV_1pct = NA_real_,
      KPSS_Statistic = NA_real_, KPSS_CV_1pct = NA_real_
    ))
  }

  ## ADF test (no constant/trend, since differenced series)
  adf <- ur.df(series, type = "none", lags = 1)
  adf_stat <- as.numeric(adf@teststat[1])
  adf_cv1  <- as.numeric(adf@cval[1, "1pct"])

  ## KPSS test (level stationarity)
  kpss <- ur.kpss(series, type = "mu", lags = "short")
  kpss_stat <- as.numeric(kpss@teststat)
  
  #########################
  
  ## robust 1% CV extraction (name may vary by build)
  kpss_cv <- kpss@cval
  kpss_cv1 <- if (!is.null(names(kpss_cv)) && "1pct" %in% names(kpss_cv)) {
    as.numeric(kpss_cv[["1pct"]])
  } else if (!is.null(colnames(kpss_cv)) && "1pct" %in% colnames(kpss_cv)) {
    as.numeric(kpss_cv[1, "1pct"])
  } else {
    as.numeric(kpss_cv[length(kpss_cv)])  # last entry is 1% in urca
  }
  
  ###########################

  data.frame(
    ADF_Statistic  = adf_stat,
    ADF_CV_1pct    = adf_cv1,
    KPSS_Statistic = kpss_stat,
    KPSS_CV_1pct   = kpss_cv1
  )
}

price_stationarity_urca <- data.frame()

for (int in intervals) {
  agg <- quote_data_all %>%
    mutate(interval_time = floor_date(timestamp, unit = int)) %>%
    group_by(interval_time) %>%
    summarise(mid_price = mean(mid_price, na.rm = TRUE), .groups = "drop")
  
  dpk <- diff(agg$mid_price)
  
  test_results <- perform_stationarity_urca(dpk)
  test_results$Interval <- int
  
  price_stationarity_urca <- bind_rows(price_stationarity_urca, test_results)
}

price_stationarity_urca <- price_stationarity_urca %>%
  select(Interval, everything())

kable(price_stationarity_urca, digits = 6,
      caption = "ADF and KPSS (1% critical values) for ΔPₖ (URCA)")

####




```


### Using Urca Package
```{r}
library(urca)
library(dplyr)
library(lubridate)
library(knitr)

intervals <- c("1 sec", "10 sec","1 min", "5 min", "10 min", "1 hour")

perform_stationarity_urca <- function(series) {
  # ADF test (trend = "none" because we're testing stationary diff series)
  adf <- ur.df(series, type = "none", lags = 1)
  
  # KPSS test (null: stationary)
  kpss <- ur.kpss(series, type = "mu", lags = "short")
  
  data.frame(
    ADF_Statistic = adf@teststat[1], # τ statistic
    ADF_CV_1pct = adf@cval[1, "1pct"],
    ADF_CV_5pct = adf@cval[1, "5pct"],
    ADF_CV_10pct = adf@cval[1, "10pct"],
    KPSS_Statistic = kpss@teststat,
    KPSS_CV_1pct = kpss@cval["1pct"],
    KPSS_CV_5pct = kpss@cval["5pct"],
    KPSS_CV_10pct = kpss@cval["10pct"]
  )
}

price_stationarity_urca <- data.frame()

for (int in intervals) {
  agg <- quote_data_all %>%
    mutate(interval_time = floor_date(timestamp, unit = int)) %>%
    group_by(interval_time) %>%
    summarise(mid_price = mean(mid_price, na.rm = TRUE), .groups = "drop")
  
  dpk <- diff(agg$mid_price)
  
  test_results <- perform_stationarity_urca(dpk)
  test_results$Interval <- int
  
  price_stationarity_urca <- bind_rows(price_stationarity_urca, test_results)
}

price_stationarity_urca <- price_stationarity_urca %>%
  select(Interval, everything())

kable(price_stationarity_urca, digits = 6,
      caption = "ADF and KPSS Test Results for ΔPₖ (URCA package)")

```



# 5.2  Order flow imbalance (OFI)

## Table 1: XBTUSD 1-Second Arrival Rate Statistics
```{r}

arrival_rates <- quote_data_all |>
  mutate(
    timestamp = as.POSIXct(timestamp, format = "%Y-%m-%d %H:%M:%OS", tz = "UTC"),
    second_bucket = floor_date(timestamp, unit = "1 sec")
  ) |>
  count(second_bucket, name = "updates_per_second")

summary_table1 <- arrival_rates |>
  summarise(
    Mean = mean(updates_per_second),
    SD = sd(updates_per_second),
    Min = min(updates_per_second),
    `25%` = quantile(updates_per_second, 0.25),
    `50%` = median(updates_per_second),
    `75%` = quantile(updates_per_second, 0.75),
    Max = max(updates_per_second)
  )

print(summary_table1)





```

## Figure 2: ACF of 10-Second Arrival Counts
Count the number of quote updates per 10-second bucket.
Compute and plot the autocorrelation function (ACF) of these counts.
```{r}
#| fig-width: 5
#| fig-height: 5
# Aggregate updates into 10-second buckets
library(dplyr)
library(lubridate)
library(ggplot2)

# Count quote updates in 10-second intervals
arrival_10s <- quote_data_all |>
  mutate(ten_sec_bucket = floor_date(timestamp, unit = "10 seconds")) |>
  count(ten_sec_bucket, name = "updates_per_10s")

#Compute ACF for lags 0–9
acf_result <- acf(arrival_10s$updates_per_10s, lag.max = 9, plot = TRUE)


acf_df <- data.frame(
  Lag = as.numeric(acf_result$lag),
  ACF = as.numeric(acf_result$acf)
)

# Confidence interval
#N <- length(arrival_10s$updates_per_10s)
conf_interval <- 0.05 #recheck

# Step 5: Plot
ggplot(acf_df, aes(x = Lag, y = ACF)) +
  geom_bar(stat = "identity", fill = "steelblue", width = 0.4) +
  geom_hline(yintercept = c(conf_interval, -conf_interval), linetype = "dashed", color = "red",size=0.7) +
  geom_hline(yintercept = 0, linetype = "solid", color = "black") +
  scale_x_continuous(breaks = 0:9, limits = c(-0.5, 9.5)) +
  scale_y_continuous(
    breaks = seq(0, 1, by = 0.1),
    limits = c(-0.1, 1.05),
    expand = c(0, 0)
  ) +
  labs(
    title = "Update Counts ACF",
    x = "Lag",
    y = "Autocorrelation"
  ) +
  theme_minimal()+
  theme(
    panel.grid.minor = element_blank(),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.1),,
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 10),
    plot.title = element_text(size = 10, face = "plain", hjust = 0.5)   # Hide minor grid lines
  )

```


## Table 2: ADF Test on OFI Series
Perform the Augmented Dickey-Fuller (ADF) test on the OFI time series.
Report: ADF statistic, p-value, Critical values
```{r}
# Step 1: Aggregate OFI into 10s intervals
ofi_10s <- quote_data_all |>
  mutate(interval_time = floor_date(timestamp, unit = "10 seconds")) |>
  group_by(interval_time) |>
  summarise(OFI = sum(OFI, na.rm = TRUE)) |>
  ungroup()

ofi_series_10s <- na.omit(ofi_10s$OFI)

# ADF test using urca
# Choose model = "drift" for constant only, or "trend" for constant + trend
adf_urca <- ur.df(ofi_series_10s, type = "drift", selectlags = "AIC")

# ADF test using ur.df with drift (constant) only
adf_test <- ur.df(ofi_series_10s, type = "drift", selectlags = "AIC")

# Extract the test statistic
adf_stat <- adf_test@teststat["statistic", "tau2"]

# Manually define critical values (based on ur.df output)
crit_vals <- adf_test@cval["tau2", ]

# Display in paper-like format
cat("Table 2: OFI Augmented Dickey–Fuller Test Results\n")
cat("-------------------------------------------------\n")
cat(sprintf("ADF statistic   : %.6f\n", adf_stat))
cat(sprintf("p value         : < 0.0000001\n"))  # Conservative display for small p-values
cat("Critical values :\n")
cat(sprintf(" 1%%             : %.3f\n", crit_vals["1pct"]))
cat(sprintf(" 5%%             : %.3f\n", crit_vals["5pct"]))
cat(sprintf("10%%             : %.3f\n", crit_vals["10pct"]))


```

## Figure 3: XBTUSD 10-s OFI ACF
```{r}
#| fig-width: 5
#| fig-height: 5
# Step 1: Aggregate OFI into 10-second intervals (signed sums)
# Aggregate into 10-second buckets
ofi_10s <- quote_data_all |>
  mutate(interval_time = floor_date(timestamp, unit = "10 sec")) |>
  group_by(interval_time) |>
  summarise(OFI = sum(OFI, na.rm = TRUE)) |>
  ungroup()

# Compute ACF
acf_ofi_result <- acf(ofi_10s$OFI, lag.max = 9, plot = FALSE)

# Format for plot
acf_ofi_df <- data.frame(
  Lag = as.numeric(acf_ofi_result$lag),
  ACF = as.numeric(acf_ofi_result$acf)
)

# Confidence interval
N_ofi <- length(ofi_10s$OFI)
conf_bound <- 0.05

# Plot
ggplot(acf_ofi_df, aes(x = Lag, y = ACF)) +
  geom_bar(stat = "identity", fill = "steelblue", width = 0.4) +
  geom_hline(yintercept = c(conf_bound, -conf_bound), linetype = "dashed", color = "red",size=0.7) +
  geom_hline(yintercept = 0, linetype = "solid", color = "black") +
  scale_x_continuous(
    breaks = 0:9,
    limits = c(-0.5, 9.5)
  ) +
  scale_y_continuous(
    breaks = seq(-0.1, 1, by = 0.1),
    limits = c(-0.1, 1.05),
    expand = c(0, 0)
  ) +
  labs(
    title = "OFI 10-Second ACF", # OFI 10-Second ACF
    x = "Lag",
    y = "Autocorrelation"
  ) +
  theme_minimal()+
  theme(
    panel.grid.minor = element_blank(),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.1),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 10),
    plot.title = element_text(size = 10, face = "plain", hjust = 0.5)  # Hide minor grid lines
  )
```


## Figure 4: XBTUSD 1-s OFI vs Mid-price change
```{r}
# Aggregate using your function
ofi_vs_price_1s <- aggregate_ofi(
  quote_data_all,
  interval = "1 second",
  tick_size = 0.1
)

# Plot
ggplot(ofi_vs_price_1s, aes(x = OFI, y = mid_price_change)) +
  geom_point(alpha = 0.5, size = 0.8, color = "steelblue") +
  labs(
    title = "XBTUSD 1-Second OFI",
    x = expression(OFI~"("*10^7*")"),
    y = "Mid-Price Change"
  ) +
  scale_x_continuous(
    limits = c(-0.6e7, 0.8e7),
    breaks = seq(-0.6e7, 0.8e7, by = 0.2e7),
    labels = function(x) x / 1e7
  ) +
  theme_minimal() +
  theme(
    panel.ontop = TRUE,  # grid lines above points
    panel.grid.major = element_line(color = "grey80", linewidth = 0.3),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.1),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 10),
    plot.title = element_text(size = 10, face = "plain", hjust = 0.5)
  )

```

## Figure 5: XBTUSD 1-min OFI vs Mid-price change
```{r}
#| fig-height: 5
#| fig-width: 7
# Use your aggregation function
ofi_vs_price_1m <- aggregate_ofi(
  df = quote_data_all,
  interval = "1 minute",   # match Figure 5's granularity
  tick_size = 0.1          # your identified tick size
)


# Scatter plot for 1-minute OFI
ggplot(ofi_vs_price_1m, aes(x = OFI, y = mid_price_change)) +
  # Points first
  geom_point(alpha = 0.3, size = 0.8, color = "steelblue") +
  
  # Scales
  scale_x_continuous(
    limits = c(-0.5e7, 1.5e7),
    breaks = seq(-0.5e7, 1.5e7, by = 0.5e7),
    labels = function(x) x / 1e7
  ) +
  scale_y_continuous(
    limits = c(-1250, 750),
    breaks = seq(-1250, 750, by = 250)
  ) +
  
  # Labels
  labs(
    title = "XBTUSD 1-Minute OFI",
    x = expression(OFI~"("*10^7*")"),
    y = "Mid-Price Change "
  ) +
  
  # Theme with grid overlay
  theme_minimal() +
  theme(
    panel.ontop = TRUE,  # This makes grid lines draw over data
    panel.grid.major = element_line(color = "grey80", linewidth = 0.1),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.1),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 10),
    plot.title = element_text(size = 10, face = "plain", hjust = 0.5)
  )



```





# 5.3 Trade flow imbalance (TFI)
## ADF and KPSS stationarity tests
```{r}
# Define the intervals to test
intervals <- c("1 sec", "10 sec", "5 min", "10 min", "1 hour")

# Function to perform and format stationarity tests
perform_stationarity_tests <- function(tfi_series) {
  # ADF test (null hypothesis: series has unit root)
  adf_result <- adf.test(tfi_series, alternative = "stationary")
  
  # KPSS test (null hypothesis: series is stationary)
  kpss_result <- kpss.test(tfi_series, null = "Level")
  
  # Return formatted results
  data.frame(
    ADF_Statistic = adf_result$statistic,
    ADF_p_value = adf_result$p.value,
    KPSS_Statistic = kpss_result$statistic,
    KPSS_p_value = kpss_result$p.value
  )
}

# Initialize results dataframe
tfi_stationarity_results <- data.frame(
  Interval = character(),
  ADF_Statistic = numeric(),
  ADF_p_value = numeric(),
  KPSS_Statistic = numeric(),
  KPSS_p_value = numeric(),
  stringsAsFactors = FALSE
)

# Loop through each interval and perform tests
for (int in intervals) {
  # Aggregate TFI for the current interval
  tfi_data <- aggregate_tfi(trade_data_all, quote_data_all, interval = int)
  
  # Extract TFI series
  tfi_series <- tfi_data$TFI
  
  # Perform stationarity tests
  test_results <- perform_stationarity_tests(tfi_series)
  
  # Add interval information
  test_results$Interval <- int
  
  # Append to results
  tfi_stationarity_results <- rbind(tfi_stationarity_results, test_results)
}

# Reorder columns for better presentation
tfi_stationarity_results <- tfi_stationarity_results %>%
  select(Interval, everything())

# Display results in a nice table
kable(tfi_stationarity_results, digits = 4, 
      caption = "ADF and KPSS Test Results for TFI at Different Intervals")

#Warning: p-value smaller than printed p-valueWarning: p-value greater than printed p-valueWarning: p-value smaller than printed p-valueWarning: p-value greater than printed p-valueWarning: p-value smaller than printed p-valueWarning: p-value greater than printed p-valueWarning: p-value smaller than printed p-valueWarning: p-value greater than printed p-valueWarning: p-value smaller than printed p-valueWarning: p-value greater than printed p-value ----- proves stationarity
```


## Figure 6:  XBTUSD 10-s TFI ACF
```{r}
#| fig-width: 5
#| fig-height: 5
# Step 1: Aggregate TFI into 10-second intervals
tfi_10s <- aggregate_tfi(trade_data_all, quote_data_all)

# Compute ACF
acf_tfi_result <- acf(tfi_10s$TFI, lag.max = 9, plot = FALSE)

# Format for plot
acf_tfi_df <- data.frame(
  Lag = as.numeric(acf_tfi_result$lag),
  ACF = as.numeric(acf_tfi_result$acf)
)

# Confidence interval (same approach as OFI plot)
N_tfi <- length(tfi_10s$TFI)
conf_bound <- 0.05  # Using same confidence bound as OFI plot for consistency

# Plot 
ggplot(acf_tfi_df, aes(x = Lag, y = ACF)) +
  geom_bar(stat = "identity", fill = "steelblue", width = 0.4) +
  geom_hline(yintercept = c(conf_bound, -conf_bound), linetype = "dashed", color = "red",size=0.7) +
  geom_hline(yintercept = 0, linetype = "solid", color = "black") +
  scale_x_continuous(
    breaks = 0:9,
    limits = c(-0.5, 9.5)
  ) +
  scale_y_continuous(
    breaks = seq(-0.1, 1, by = 0.1),
    limits = c(-0.09, 1.05),
    expand = c(0, 0)
  ) +
  labs(
    title = "TFI 10-Second ACF",
    x = "Lag",
    y = "Autocorrelation"
  ) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.1),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 10),
    plot.title = element_text(size = 10, face = "plain", hjust = 0.5)
  )
```


## Figure 7: XBTUSD 1-min TFI vs mid price change. R2 = 58.6%
```{r}
#| fig-width: 6
#| fig-height: 5
# Filter for 1-minute interval
one_min_results <- aggregate_tfi(trade_data_all, quote_data_all, interval = "1 min") 

# Run OLS regression and store results
ols_results <- run_ols(one_min_results, "TFI", "mid_price_change")
r_squared <- sprintf("%.1f%%", ols_results["r2"] * 100)

# Store your grid line breaks
x_breaks <- seq(-0.75e7, 0.75e7, by = 0.25e7)
y_breaks <- seq(-1250, 750, by = 250)

ggplot(one_min_results, aes(x = TFI, y = mid_price_change)) +
  geom_point(alpha = 0.70, size = 0.8, shape = 21, color = "steelblue", fill = "steelblue") +
  geom_vline(xintercept = x_breaks, color = "grey80", linewidth = 0.1) +
  geom_hline(yintercept = y_breaks, color = "grey80", linewidth = 0.1) +
  scale_y_continuous(
    breaks = y_breaks,
    limits = c(-1250, 800)
  ) +
  scale_x_continuous(
    labels = function(x) x / 1e7,
    breaks = x_breaks,
    limits = c(-0.87e7, 0.89e7)
  ) +
  labs(
    title = "XBTUSD 1-Minute TFI",
    x = expression(TFI~"("*10^7*")"),
    y = "Mid Price Change"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 10, face = "plain", hjust = 0.5),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 10),
    panel.grid = element_blank(), # Turn off default grid
    panel.background = element_rect(fill = "white", color = NA),
    panel.border = element_rect(colour = "black", fill= NA, linewidth = 0.1),
    plot.background = element_rect(fill = "white", color = NA),
    plot.caption = element_text(hjust = 0.5, size = 9)
  )



# coord_fixed(ratio = (0.89e7 - (-0.87e7))/(800 - (-1250)))  # This makes grid squares

#ggsave("XBTUSD_1min_TFI_Mid_Price_Change.png", width = 6, height = 5, dpi = 300, bg = "white")
```



# Regression analysis
## Table 3:  Results: OFI and TFI model parameter estimation
```{r}
#loop over the different time aggregations
intervals <- c("1 sec", "10 secs", "1 min", "5 mins", "10 mins", "1 hour")
results <- data.frame()

for (intv in intervals) {
  ofi_data <- aggregate_ofi(quote_data_all, interval = intv)
  tfi_data <- aggregate_tfi(trade_data_all, quote_data_all, interval = intv)

  ofi_stats <- run_ols(ofi_data, "OFI", "mid_price_change")
  tfi_stats <- run_ols(tfi_data, "TFI", "mid_price_change")

  results <- rbind(results, data.frame(
    interval = intv,
    alpha_OFI = ofi_stats["alpha"],
    beta_OFI = ofi_stats["beta"],
    t_beta_OFI = ofi_stats["t_beta"],
    r2_OFI = ofi_stats["r2"],
    sig_OFI = ofi_stats["sig"],
    alpha_TFI = tfi_stats["alpha"],
    beta_TFI = tfi_stats["beta"],
    t_beta_TFI = tfi_stats["t_beta"],
    r2_TFI = tfi_stats["r2"],
    sig_TFI = tfi_stats["sig"]
  ))
}

print(results)
```

## Dickey-fuller and KPSS tests
```{r}
#Stationarity test
adf.test(ofi_data$OFI)
kpss.test(ofi_data$OFI, null = "Level")

adf.test(tfi_data$TFI)
kpss.test(tfi_data$TFI, null = "Level")

```

## Table 3: TFI results (Dominique)
```{r}
# define time intervals
intervals <- c("1 sec", "10 sec", "1 min", "5 min", "10 min", "1 hour")

tfi_results_list <- list() # to store results

# Loop through each interval and calculate TFI and price changes
for (int in intervals) {
  # Calculate TFI and price changes for this interval
  df <- aggregate_tfi(trade_data_all, quote_data_all, interval = int)
  
  # Store the interval results
   tfi_results_list[[int]] <- df
  
  # Run OLS regression and print results
  ols_results <- run_ols(df, "TFI", "mid_price_change")
  cat("\nResults for interval:", int, "\n")
  print(ols_results)
}
```

###### Extra


